Perturbing SU($n$)-symmetric interactions

### 2019-10-30 collective_sun/sun_perturbation.tex
added section on finding eigenvectors of the non-uniform interaction Hamiltonian

\renewcommand{\d}{\text{d}} % for infinitesimals
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

@@

depend on $\v v$.  Finding vectors $\v v$ that satisfy
\eqref{eq:v_cond_1} is equivalent to finding eigenvectors of the
matrix $\m h-\m h_\d$, where $\m h_\d\equiv\diag\v h_\d$ is a diagonal
matrix; the constant $\Delta\p{\v v}$ is then the eigenvalue of
$\m h-\m h_\d$ corresponding to the eigenvector $\v v$.  If the
interaction Hamiltonian $H_0$ is translationally invariant, then all
$h_q=h$ and $\v h_\d\oc\v v=h\v v$, so the condition in
\eqref{eq:v_cond_1} simplifies to

@@

satisfies, for some constant $\varepsilon\p{\m v}$ that may depend on
$\m v$,

@@

  \tilde{\m h}\oc\m v + \m h\c\m v
  = \varepsilon\p{\m v} \m v,

@@

  \tilde{\m h}

@@

where $\tilde{\m h}\oc\m v$ denotes an element-wise (Hadamard) product
of $\tilde{\m h}$ and $\m v$, with components

@@

\eqref{eq:H_V2_psi_mid} vanishes.  The condition in
\eqref{eq:v_cond_2} can be recast as an eigenvalue equation by
treating the $N\times N$ matrix $\m v$ as a vector with $N^2$
components, which can be further reduced to ${N\choose2}=N\p{N-1}/2$
components by the fact that the matrix $\m v$ is symmetric with zeros
on the diagonal.  If the interaction Hamiltonian $H_0$ is

@@

  \m h\oc\m v + \m h\c\m v = \sp{\varepsilon\p{\m v}+h} \m v,

@@

  = \sp{E_0 + 2\varepsilon\p{\m v}} \V_2\p{\m v} \ket\psi.

@@

  N \varepsilon\p{\m v} v = \f12 \sum_{k,\ell} h_{k\ell} v_{k\ell},

@@

>>>>

\renewcommand{\d}{\text{d}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\e}{\varepsilon}

@@

depend on $\v v$.  If the interaction Hamiltonian $H_0$ is
$\v h_\d\oc\v v=h\v v$, so the condition in \eqref{eq:v_cond_1}
simplifies to

@@

satisfies, for some constant $\e\p{\m v}$ that may depend on $\m v$,

@@

  \bar{\m h}\oc\m v + \m h\c\m v = \e\p{\m v} \m v,

@@

  \bar{\m h}

@@

where $\bar{\m h}\oc\m v$ denotes an element-wise (Hadamard) product
of $\bar{\m h}$ and $\m v$, with components

@@

\eqref{eq:H_V2_psi_mid} vanishes.  If the interaction Hamiltonian
$H_0$ is translationally invariant, then all $h_q=h$ and

@@

  \m h\oc\m v + \m h\c\m v = \sp{\e\p{\m v}+h} \m v,

@@

  = \sp{E_0 + 2\e\p{\m v}} \V_2\p{\m v} \ket\psi.

@@

  N \e\p{\m v} v = \f12 \sum_{k,\ell} h_{k\ell} v_{k\ell},

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finding eigenstates of the interaction Hamiltonian}

Here we discuss the construction of eigenstates of the interaction
Hamiltonian $H_0$.  If a vector $\v v$ and matrix $\m v$ satisfy the
conditions in \eqref{eq:v_cond_1} and \eqref{eq:v_cond_2}, then the
single- and two-body operators $\V_1\p{\v v}$ and $\V_2\p{\m v}$
generate eigenvectors of the interaction Hamiltonian $H_0$ when they
act on any fully symmetric state $\ket\psi\in\M_0$.  We can use this
fact to construct eigenstates of $H_0$ by finding solutions to
\eqref{eq:v_cond_1} and \eqref{eq:v_cond_2}.

The single-body case is straightforward: finding vectors $\v v$ that
satisfy \eqref{eq:v_cond_1} is equivalent to finding eigenvectors of
the matrix $\m h-\m h_\d$, where $\m h_\d\equiv\diag\v h_\d$ is a
diagonal matrix.  The constant $\Delta\p{\v v}$ is then the eigenvalue
of $\m h-\m h_\d$ that corresponds to the eigenvector $\v v$.  The
task of constructing eigenvectors $\V_1\p{\v v}\ket\psi$ of $H_0$ and
determining their energy thus reduces to finding the eigenvectors and
eigenvalues of the $N\times N$ matrix $\m h-\m h_\d$.

The two-body case requires a bit of work to re-cast as an
eigenvector/eigenvalue problem.  We begin by writing the matrix $\m v$ as a vector,
\begin{align}
  \m v = \sum_{p,q} v_{pq} \op{k}{q}
  \to \check{\m v} \equiv \sum_{p,q} v_{pq} \ket{pq},
\end{align}
and then defining the matrix
\begin{align}
  \check{\m h}
  \equiv \sum_{k,\ell} \bar h_{k\ell} \op{k\ell}
  + \sum_{k,q,\ell} h_{kq} \op{k}{q} \otimes \op{\ell},
  \label{eq:h_super_mat}
\end{align}
in terms of which the condition in \eqref{eq:v_cond_2} becomes
\begin{align}
  \check{\m h} \c \check{\m v} = \e\p{\m v} \check{\m v}.
\end{align}
Note that the matrix $\check{\m h}$ in \eqref{eq:h_super_mat} is
block-diagonal, with $N$ blocks of size $N\times N$.  The task of
constructing eigenvectors $\V_2\p{\m v}\ket\psi$ of $H_0$ and
determining their energy thus reduces to finding the eigenvectors and
eigenvalues of $N$ matrices of size $N\times N$.
